\section{Discussion}
\label{sec:discussion}

\para{The hypothesis holds, with one decisive correction.} Misalignment
propagation \emph{does} obey epidemiological dynamics. We measured a real
per-contact transmission probability and a recovery process, and the calibrated
network model produced every classical signature: a basic reproduction number, a
sharp spectral phase transition at $R_0=1$, topology-dependent vulnerability, and
critical intervention thresholds. The correction is that the load-bearing
parameter is per-model \emph{susceptibility}, not topology. Our measurement shows
susceptibility is binary-like and model-specific---\gptmini is immune while
\llama is susceptible---and in the SIS/SIR framework this is exactly the immunity
term. The herd-immunity sweep shows a sufficiently robust population
(${\sim}70$--$80\%$ immune) cannot sustain an epidemic regardless of topology.
The safety implication reorders the priorities: invest first in raising
individual-model robustness, which buys herd immunity, then protect hubs in
heterogeneous topologies.

\para{Channel and payload matter.} The narrow insecure-code payload that reliably
induces EM under \emph{fine-tuning} did not transmit broad misalignment
\emph{in-context}. This is consistent with the persona-features account
\citep{wang2025persona}: broad EM is mediated by an activated ``misaligned
persona'' that is elicited far more strongly by demonstrations of misaligned
\emph{behavior} (E1b) than by narrow code artifacts (E1a). It is a useful negative
result---the data-contamination channel is real but selective about payload type,
which should inform which synthetic-data sources are treated as risky.

\para{Relation to prior results.} The spectral threshold and the hub-targeting
advantage reproduce canonical network-epidemic results
\citep{pastorsatorras2015epidemic,castellano2010thresholds}. The measured
susceptible-model transmission (${\sim}0.16$) coincidentally matches the
literature prior, and the steep low-dose onset echoes
\citet{qi2023finetuning}. Fast, complete recovery echoes the efficacy of
re-alignment via safety vectors~\citep{bhardwaj2024resta}, while the immune model
is the benign analog of the persistent carrier studied by
\citet{hubinger2024sleeper}---in our measurements recovery is large, but a
backdoor carrier with $\mu\approx 0$ would change the picture.

\para{Limitations.} We are explicit about scope.
\emph{(1) In-context $\neq$ weight merging.} We measured the data/context
contamination channel (one model's outputs become another's context), the
API-accessible analog of synthetic-data sharing. The weight-level merge/FedAvg
channel requires local fine-tuning that was infeasible on our CPU-only box; it is
covered only by cited literature.
\emph{(2) Two model families.} Immune-versus-susceptible is established only on
\gptmini and \llama; the true immune \emph{fraction} in a real ecosystem is
unknown, so we sweep it rather than estimate it.
\emph{(3) Modest samples.} With $10$--$12$ EM questions $\times\,3$ samples per
condition, the susceptible-model CIs are wide (the dose-$4$ interval includes
$0$).
\emph{(4) Dose-to-round mapping.} Translating in-context dose into a per-round
contact probability is a modeling choice; we mitigate by sweeping $\beta/\mu$
across the full phase diagram and marking the measured point.
\emph{(5) Judge monoculture.} A single judge (\gptmini) scores alignment;
although this is the standard EM protocol, judge bias is uncontrolled.
\emph{(6) Static, homogeneous networks.} We use static topologies and a
homogeneous $\beta$ within the susceptible class; switching networks and per-node
heterogeneity are left to future work.

\para{Broader impact.} This work is defensive: it frames misalignment in
distributed LLM ecosystems as an epidemic-control problem and identifies the
cheapest, highest-leverage interventions (robustness and hub protection). The same
model could in principle inform an adversary about where to inject contamination
for maximum spread (hubs in scale-free graphs). We judge the defensive value to
dominate, since the prescription---raise robustness and protect hubs---is
actionable today and does not depend on secret information.
